---
title: "Week05"
author: "Rebecca"
date: "4/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(tidycensus)
```
```{r}
#source(here::here("Scripts", "get_hud_data.R"))
```


```{r}

hud<- read_csv(here::here("Data_Transformation", "hud_2020_tract_joined.csv"))
```
```{r}
glimpse(hud)
```
```{r}
# some of the logicals may not be logical
summary(hud)

```
# some efiels are all NA
# some must be amany missing

```{r}

hud%>% 
  select (state, entities, place, cbsa, code, program_label, program, everything(), 
          -where(is.logical), -gsl, -sumlevel, -Quarter, -states, - program) %>% 
  mutate(across(where(is.double(), ~if_else(. == -1, NA_real_, .)))) %>% 
  filter(state == "DE", 
          program_label == "Housing Choice Vouchers")
```
```{r}

hud_de <- 
  hud %>% 
  select(state, entities, place, cbsa, code, program_label, program, name, everything(), 
         -where(is.logical), -gsl, -sumlevel, -Quarter, -states, -program) %>% 
  mutate(across(where(is.double), ~ if_else(. == -1, NA_real_, .))) %>% 
  filter(state == "DE",
         program_label == "Housing Choice Vouchers")
```

```{r}
hud_de %>% 
  ggplot(aes (x = total_units)) +
  geom_histogram()



```
```{r}
# we need long data instead of wide data

#note there is already a field called name
hud_de %>% 
  pivot_longer(where(is.double)
               , names_to = "col") %>% 
  mutate(value = if_else(value == -4, NA_real_, value)) %>% 
    ggplot(aes (x = value)) +
  geom_histogram() +
  facet_wrap(~col, scales = "free_x")


# need to adjust for the -4
```
```{r}

```
#american commintiy data
#census data
```{r}
library(tidycensus)

```
in console
census_api_key("YOUR API KEY GOES HERE", install = TRUE)
# be sure these are the right quotes!!!

```{r}
readRenviron("~/.Renviron")


```

```{r}
load_variables(2018, "acs5")


```

# name is a variable code for the variebles we want


```{r}
get_acs(geography = "tract",
        variables = c(total = "B00001_001") ,
        state = "DE")
```


```{r}
get_acs(geography = "tract",
        variables = c(total = "B01001_001",
                      male = "B01001_002"),
        state = "DE")


```
```{r}
de_gender <- 
get_acs(geography = "tract",
        variables = c(total = "B01001_001",
                      male = "B01001_002"),
        state = "DE", 
        output = "wide")


```
```{r}
# E is estimate
# M is margin of error

de_gender %>% 
  mutate(male_percent = maleE / totalE) %>% 
  arrange(desc(male_percent))
```

```{r}
pop_estimates <- c(
  pe.med_inc             = 'B06011_001',    # median income
  pe.med_f               = 'B10010_001',    # median family income
  pe.hh_inc              = 'B19001_001',    # household income in the past 12 months
  pe.tot_hh              = 'B11001_001',    # total households
  pe.renter_householders = 'B07013_003',    # total renter households
  pe.tot_pop_rent        = 'B25008_003',    # totol population in renter occupied homes
  pe.tot_hu              = 'B25001_001',    # total housing units
  pe.pop_hu              = 'B25008_001',    # total population in occupied housing units
  pe.popTotE             = 'B01003_001'     # total population
)


# Select the columns corresponding to the counts of hte
# households who are paying 30 percent or more on rent
# https://censusreporter.org/data/table/?table=B25070&geo_ids=140%7C04000US10&primary_geo_id=04000US10#valueType%7Cestimate
income_percent <- c(
  inc_per.rent_00 = 'B25070_002',    # households paying less than 10.0 percent in rent
  inc_per.rent_10 = 'B25070_003',    # households paying 10.0 to 14.9 percent in rent
  inc_per.rent_15 = 'B25070_004',    # households paying 15.0 to 19.9 percent in rent
  inc_per.rent_20 = 'B25070_005',    # households paying 20.0 to 24.9 percent in rent
  inc_per.rent_25 = 'B25070_006',    # households paying 25.0 to 29.9 percent in rent
  inc_per.rent_30 = 'B25070_007',    # households paying 30.0 to 34.9 percent in rent
  inc_per.rent_35 = 'B25070_008',    # households paying 35.0 to 39.9 percent in rent
  inc_per.rent_40 = 'B25070_009',    # households paying 40.0 to 49.9 percent in rent
  inc_per.rent_50 = 'B25070_010'     # households paying 50.0 percent or more in rent
)

# Get Rent Categories x Household Income Table (B25074)
# For each income category and percentage rent spend on household income:
# (10k = less than 10k, 20k = less than 20k, 35k = less than 35k, 50k = less than, and 75k = less than 75k)
percent_by_income <- c(
  income.rent_30_10k = 'B25074_006',
  income.rent_35_10k = 'B25074_007',
  income.rent_40_10k = 'B25074_008',
  income.rent_50_10k = 'B25074_009',
  income.rent_30_20k = 'B25074_015',
  income.rent_35_20k = 'B25074_016',
  income.rent_40_20k = 'B25074_017',
  income.rent_50_20k = 'B25074_018',
  income.rent_30_35k = 'B25074_024',
  income.rent_35_35k = 'B25074_025',
  income.rent_40_35k = 'B25074_026',
  income.rent_50_35k = 'B25074_027',
  income.rent_30_50k = 'B25074_033',
  income.rent_35_50k = 'B25074_034',
  income.rent_40_50k = 'B25074_035',
  income.rent_50_50k = 'B25074_036',
  income.rent_30_75k = 'B25074_042',
  income.rent_35_75k = 'B25074_043',
  income.rent_40_75k = 'B25074_044',
  income.rent_50_75k = 'B25074_045'
)

cost_by_size <- c(
  cost.med_rent  = 'B25031_001',    # median gross rent (total)
  cost.0_bed     = 'B25031_002',    # median gross rent (no bedroom)
  cost.1_bed     = 'B25031_003',    # median gross rent (one bedroom)
  cost.2_bed     = 'B25031_004',    # median gross rent (two bedrooms)
  cost.3_bed     = 'B25031_005',    # median gross rent (three bedrooms)
  cost.4_bed     = 'B25031_006',    # median gross rent (four bedrooms)
  cost.5_bed     = 'B25031_007'     # median gross rent (five bedrooms)
)
```



```{r}
allcensus_vars <- c( pop_estimates,
                     income_percent,
                     percent_by_income,
                     cost_by_size)
```
```{r}
acs_df <- get_acs(
  geography = "tract",
  variables = allcensus_vars,
  state = "DE"
)
  

```
```{r}
acs_df
```




```{r}
acs_df <- 
  acs_df %>% 
  select(-moe)
```


```{r}
acs_df %>% 
  filter(str_starts(variable, "cost")) %>% 
  ggplot( aes(x=estimate)) +
  geom_histogram() +
  facet_wrap(~variable)
```



```{r}
acs_df %>% 
  filter(str_starts(variable, "income")) %>% 
  ggplot( aes(x=estimate)) +
  geom_histogram() +
  facet_wrap(~variable)


```



```{r}
acs_df %>% 
  filter(str_starts(variable, "inc_per")) %>% 
  ggplot( aes(x=estimate)) +
  geom_histogram() +
  facet_wrap(~ variable)
```


```{r}
acs_df %>% 
  filter(str_starts(variable, "")) %>% 
  ggplot( aes(x=estimate)) +
  geom_histogram() +
  facet_wrap(~variable)
```

#combining the data
```{r}
#we want 1 row for every census tract
#we want to combine both data sets
acs_df_wide <- 
  acs_df %>% 
    pivot_wider(names_from = variable, 
                values_from = estimate)


```

#join
```{r}
#left_join(x, y, by)

left_join(
  x = acs_df_wide ,
  y = hud_de ,
  by = c("GEOID" = "code")
)

```

```{r}
acs_df_wide %>% 
left_join(hud_de,
           by ("GEOID" = "code"))




```

```{r}
left_join(
  x = hud_de  ,
  y =  acs_df_wide,
  by = c("code" = "GEOID")
)



```
# with piping
```{r}

hud_de %>% 
  left_join( 
    y = acs_df_wide ,
      by = c("code" = "GEOID")
)

```
# if anything does not match, default is to fill with null


```{r}
housing <-
  acs_df_wide %>% 
  left_join(hud_de,
           by= c("GEOID" = "code"))

```

How does the number of reported households to use section 8 covouchers relate to the totla number of households in a community?
```{r}
housing %>% 
  mutate(number_reported = if_else(number_reported %in% c(-5, -4, -1), NA_real_, number_reported)) %>% 
  ggplot( aes( x = pe.tot_hh,
               y = number_reported)) +
  geom_point()


```


```{r}
housing %>% 
  select(GEOID, starts_with("cost"), people_per_unit) %>% 
    mutate(people_per_unit = if_else(people_per_unit %in% c(-5, -4, -1), NA_real_, people_per_unit)) %>% 
    pivot_longer(starts_with("cost"),
                 names_to = "type") %>% 
  filter(type != "cost.med_rent") %>% 
  ggplot (aes(x= people_per_unit,
              y = value)) +
  geom_point() +
  facet_wrap(~ type)
  
  
```

